What's hot and what's not:\\Windowed developer topic analysis

### 2009-04-12 lda-paper.tex
Minor edits

We now define some terms that we will use in the rest of the paper: A
\emph{message} is a block of text written by developers. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when the user commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in a message, a word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is a one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project.

@@

Our technique differs from the previous techniques because we apply
LDA locally to month-long windows of commit log comments, whereas other
approaches apply LDA once to the entire project. This allows us to
examine the time-based windowing of topics over its development
history.

@@

The query optimizer was a topic that recurred during MaxDB's
development. In our plots, topics that mention \emph{optimizer} occur
four times, yet in the global-trend-view (Figure
\ref{fig:statictopics}) it is not in any of the topics. A query
optimizer is an important component of an DBMS, but as we have shown
it does not appear as a topic on its own. We tried to remove words to
see if we could get an \emph{optimizer} topic. After removing stop
words and then two of the most common words, the global analysis
finally had a topic with optimizer in its topic ten words. This shows
that optimizer was important but it had been obscured, but would have
been noticed using the more local topic analysis.

We noticed that commits that mentioned \emph{PHP} occurred two thirds
less frequently than commits that mentioned \emph{Perl}, but
\emph{Perl}-related topics appeared in the global static topics for
MaxDB while \emph{PHP}-related topics did not.  Our local topic
analysis mentioned \emph{PHP} in 5 different topics, yet only
mentioned \emph{Perl} in 4 different topics and 1 global
topic. Perhaps this is because there was a cluster of \emph{Perl}
mentions during one month while the \emph{PHP} mentions were more
spread out.

@@

We presented several visualization techniques that focused on
different aspects of the data: temporality of trends, trend size, and
a compact-trend-view. The compact-trend-view shows much more
information than the views that global analysis could show, it
indicates how focused a period is by the total number of topics, as
well, it shows topics by similarity so one can track trends across
time.  Our trend-histogram immediately highlights and measures the
size of trends while our trend-time-line view shows how a topic
reoccurs over time. These visualizations help us get a general feeling
about the common and unique topics that developers focus on during
development. If implemented interactively a user could easily zoom in
and query for a summary of a topic or trend.

>>>>

We now define some terms that we will use in the rest of the paper: A
\emph{message} is a block of text written by developers. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when the user commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in a message, a word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project.

@@

Our technique differs from the previous techniques because we apply
LDA locally to month-long windows of commit log comments, whereas
other approaches apply LDA once to the entire project. Windowed topic
analysis allows us to examine the time-based windowing of topics over
its development history.

@@

The query optimizer was a topic that recurred during MaxDB's
development. In our plots, topics that mention \emph{optimizer} occur
four times, yet in the global-trend-view (Figure
\ref{fig:statictopics}) it is not in any of the topics. A query
optimizer is an important component of an DBMS, but as we have shown
it does not appear as a topic on its own. We tried to remove words to
see if we could get an \emph{optimizer} topic. After removing stop
words and then two of the most common words, the global analysis
finally had a topic with optimizer in its topic ten words. Our analysis shows
that optimizer was important but it had been obscured, but would have
been noticed using the more local topic analysis.

We noticed that commits that mentioned \emph{PHP} occurred two thirds
less frequently than commits that mentioned \emph{Perl}, but
\emph{Perl}-related topics appeared in the global static topics for
MaxDB while \emph{PHP}-related topics did not.  Our local topic
analysis mentioned \emph{PHP} in 5 different topics, yet only
mentioned \emph{Perl} in four different topics and one global
topic. Perhaps this is because there was a cluster of \emph{Perl}
mentions during one month while the \emph{PHP} mentions were more
spread out.

@@

We presented several visualization techniques that focused on
different aspects of the data: temporality of trends, trend size, and
a compact-trend-view. The compact-trend-view shows much more
information than the views that global analysis could show, it
indicates how focused a period is by the total number of topics, as
well, it shows topics by similarity so one can track trends across
time.  Our trend-histogram immediately highlights and measures the
size of trends while our trend-time-line view shows how a topic
reoccurs over time. These visualizations help us get a general feeling
about the common and unique topics that developers focus on during
development. If implemented interactively an user could easily zoom in
and query for a summary of a topic or trend.
